---
title: "Data Exploration"
author: "Hazim Fitri"
date: "2024-12-29"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
knitr:
  opts_chunk:
    echo: true
    include: true
    message: false
    warning: false
---

# Import data into R

## .txt file

```{r}
read.table('Rclass.txt', sep=';')
```

## .csv file

```{r}
#read.csv('Rclass.csv', sheet='Rxls')
```

## .xlsx file

```{r}
library(readxl)
#read_excel('Rclass.xlsx')
```

## Clipboard

```{r}
#read.table(file='clipboard', header=T)
```

# Export data from R

## .txt file

```{r}
mydata = data.frame()
#fix(mydata)
mydata
```

```{r}
write.table(mydata, file='Rdata.txt', col.names=T, row.names=F)
getwd()
```

```{r}

```

## .csv file

```{r}
write.csv(mydata, file='Rdata.csv', col.names=T, row.names = F)
```

## Clipboard

```{r}
write.table(mydata, col.names=T, row.names = F, file=)
```

# Descriptive Statistics

1.  Central tendencies
2.  Quartiles
3.  Variation
4.  Shape

## Central Tendencies

```{r}
head(trees)
```

```{r}
mean(trees$Girth)
rowMeans(trees)
colMeans(trees)
median(trees$Girth)
```

```{r}
library(DescTools)
Mode(trees$Girth)
```

```{r}
x <- c(1,2,2,2,3,4,5,6,7,8,9)
table(x)
Mode(x)
```

## Quartile

```{r}
# by default, it'll give the 0, 0.25, 0.5, 0.75, 100
quantile(trees$Girth,c(0.25, 0.75, 0.99))
```

```{r}
summary(trees$Girth)
```

## Measures of Variation / Scale of Parameters

```{r}
max(trees$Girth) - min(trees$Girth)
range(trees$Girth)
IQR(trees$Girth)
```

## Variance

It'll provide the sample variance and not the population variance

```{r}
#var()
#sd() # This will also provide the sample standard deviation
```

# Shape

## Skewness

```{r}
library(e1071)
```

## Kurtosis

# Exercise

```{r}
batting = read.table('batting.history.txt', header = T)
colnames(batting)

BATTING = data.frame(
  Year = batting$Year,
  Tms = batting$Tms,
  N.Bat = batting$N.Bat,
  BatAge = batting$BatAge,
  Avg_PA = mean(batting$PA),
  Avg_AB = mean(batting$AB),
  Avg_H = mean(batting$H)
)

write.csv(BATTING, 'BATTING.csv', row.names = F)
```

```{r}
new = rowMeans(batting[,7:9])
battingnew = data.frame(batting[,1:4], new)
write.csv(battingnew, file='BATTING2.csv', col.names=T, row.names=F)
```

# Data Wrangling

```{r}
#head(trees,2)
#tail(trees,2)
#tail(tees, -5) # All observation except the last 5
```

```{r}
which(mtcars$cyl  == 6)
mtcars[which(mtcars$cyl  == c(6,8)),]
```

```{r}
subset(mtcars, subset = mpg<20, select=c(mpg, disp))
```

```{r}
sort(mtcars$mpg, decreasing = T)
```

```{r}
order(mtcars$mpg)
```

```{r}
 mons <- c("March","April","January","November","January","September",
 "October","September", "November","August","January",
 "November", "November", "February","May","August","July",
 "December","August","August","September","November",
 "February","April")
 table(mons)
 
mons <- factor(mons,levels=c("January","February","March",
"April","May","June","July", "August","September", "October", 
"November", "December"), ordered=TRUE)
table(mons)

names(table(mons))
```

```{r}
lapply(trees, FUN=median)
```

```{r}
do.call(rbind, lapply(trees, FUN=median))
```

```{r}
mine<-c("a","b","c")
paste(mine, "!")
paste0(mine, "!")
paste(mine, 1:3)
paste(mine, 1:3, sep=',')
paste0(mine, 1:3, collapse=',')
```

# Exercise

## Question 2

```{r}
length(mtcars$cyl)
table(factor(mtcars$cyl))/32*100
```

```{r}
summary(mtcars$mpg)
```

```{r}
data(mtcars)

variables <- c("mpg", "disp", "hp", "drat", "wt", "qsec")

iqr_values <- sapply(mtcars[variables], IQR)

print(iqr_values)
```

```{r}
data(mtcars)

mtcars$mpg_factor <- cut(mtcars$mpg, 
                         breaks = quantile(mtcars$mpg, probs = seq(0, 1, 0.25)), 
                         include.lowest = TRUE, 
                         labels = c('Low', "Medium", "High", "Very High"))

mtcars$disp_factor <- cut(mtcars$disp, 
                          breaks = quantile(mtcars$disp, probs = seq(0, 1, 0.25)), 
                          include.lowest = TRUE, 
                          labels = c("S", "M", "L", "XL"))

factor_combination <- table(mtcars$mpg_factor, mtcars$disp_factor)

print(factor_combination)
```

## Question 3

```{r}
animal = data.frame(Farm = c('MO', 'MO', 'MO', 'MO', 'LN', 'SE', 'QM'), 
           Month = c('11', '07', '07', NA, '09', '09', '11'),
           Year = c('00', '00', '01', NA, '03', '03', '02'),
           Sex = c(1,2,2,2,1,2,2),
           LengthClass = c(1,1,1,1,1,1,1),
           LengthCT = c(75, 85, 91.6, 95, NA, 105.5, 106),
           Ecervi = c(0,0,0,NA,0,0,0),
           Tb = c(0,0,1,NA,0,0,0))

animal
```

```{r}
mean(animal$LengthCT, na.rm = T)
```

```{r}
nrow(subset(animal, animal$Sex==1))
```

```{r}
nrow(subset(animal, animal$Tb==1))
```

```{r}
library(dplyr)
animal %>% mutate(Sqrt = sqrt(animal$LengthCT))
```

```{r}
mean_animal = mean(animal$LengthCT, na.rm=T)
mae = mean(abs(animal$LengthCT - mean_animal), na.rm = T)
mae
```
