---
title: "Data Analysis with R"
author: "Hazim Fitri"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
---

# Hypothesis Testing

```{r}
library(TeachingDemos) # for z.test and t.test
library(lattice) # for ANOVA
```

## Z-test

Perform a test for a population mean with known variance.

-   **Null Hypothesis (H₀)**: The sample mean is equal to the population mean (μ = μ₀).

-   **Alternative Hypothesis (H₁)**: The sample mean is not equal to the population mean (μ ≠ μ₀).

    -   *Two-tailed*: μ ≠ μ₀

    -   *One-tailed (right)*: μ \> μ₀

    -   *One-tailed (left)*: μ \< μ₀

```{r}
x = rnorm(100, mean = 2, sd = 0.5)
z.test(x, mu = 2, stdev = 0.5, alternative = 'two.sided')
```

## One Sample t-test

Perform a test for a population mean of a small (normally distributed) sample.

-   **Null Hypothesis (H₀)**: The sample mean is equal to the population mean (μ = μ₀).

-   **Alternative Hypothesis (H₁)**: The sample mean is not equal to the population mean (μ ≠ μ₀).

    -   *Two-tailed*: μ ≠ μ₀

    -   *One-tailed (right)*: μ \> μ₀

    -   *One-tailed (left)*: μ \< μ₀

```{r}
x = rnorm(20, mean = 1, sd = 0.5)
t.test(x, mu = 2, alternative = 'less')
```

## Two Sample t-test

Perform a test for the difference of two population means that is **independent** to each other with small (normally distributed) samples

-   **H₀**: The means of the two independent groups are equal (μ₁ = μ₂).

-   **H₁**: The means of the two groups are not equal (μ₁ ≠ μ₂).

    -   *Two-tailed*: μ₁ ≠ μ₂

    -   *One-tailed*: μ₁ \> μ₂ or μ₁ \< μ₂

```{r}
x = c(1.3,1.5,1.2,1.7,1.3)
y = c(1.6,1.7,1.8,1.6,1.5)
t.test(x,y,mu=0,alternative="less",var.equal=TRUE)
```

## Paired t-test

Perform a test for the difference of two population means for **two paired samples** (differences normally distributed.

-   **H₀**: The mean difference between paired samples is zero (μd = 0).

-   **H₁**: The mean difference between paired samples is not zero (μd ≠ 0).

    -   *Two-tailed*: μd ≠ 0

    -   *One-tailed*: μd \> 0 or μd \< 0

```{r}
x = c(1.3,1.5, 1.2, 1.7, 1.3)
y = c(1.6,1.7,1.8,1.6,1.5)
t.test(x, y, mu = 0, alternative = 'less', var.equal = TRUE, paired = TRUE)
```

## ANOVA

1.  A test used to determine differences between research results from three or more unrelated samples/groups.

2.  Want to study the effect of one or more qualitative variable on a quantitative outcome.

3.  Qualitative variables -\> factors.

4.  Assumptions;

    -   Normality

    -   Equal Variance

    -   Independent Samples

$$
H_0:\mu_1=\mu_2=~...~=\mu_k
$$

$$
H_1:\mu_i\not=\mu_j,~(for~some~i~and~j)
$$

# One-way ANOVA

-   **H₀**: All group means are equal (μ₁ = μ₂ = μ₃ = ... = μk).

-   **H₁**: At least one group mean is different from the others.

```{r}
one.way = aov(yield ~ variety*site, data = barley)
summary(one.way)
```

## Two-way ANOVA

-   **H₀ (Main Effects)**: There is no difference in means for each factor.

-   **H₀ (Interaction Effect)**: There is no interaction between the two factors.

-   **H₁**: At least one factor or interaction has a significant effect.

```{r}
two.way = aov(yield ~ variety+site, data = barley)
summary(two.way)
```

## Two-way ANOVA with interaction effect

```{r}
interaction = aov(yield ~ variety * site, data = barley)
summary(interaction)
```

## One Sample Test for Proportions

Test for a population proportion for a large sample

prop.test(x, n, p, alternative = c('two.sided', 'less', 'greater'), conf.level, correct)

x = number of success

n = number of trials

p = sucess probability; default p = NULL

-   **H₀**: The population proportion is equal to a specified value (p = p₀).

-   **H₁**: The population proportion is not equal to the specified value (p ≠ p₀).

```{r}
prop.test(52, 100, p = 0.4, alternative = 'greater', correct = F)
```

## Two Sample Test for Proportions

Test for a difference in population proportions for two independent large samples

```{r}
prop.test(c(132,135), c(400, 390), alternative = 'two.sided', correct = F)
```

## Chi-Square Test for Variance

**Test for Independence**:

-   **H₀**: The two variables are independent (no association).

-   **H₁**: The two variables are not independent (there is an association).

**Test for Goodness of Fit**:

-   **H₀**: The observed frequencies fit the expected distribution.

-   **H₁**: The observed frequencies do not fit the expected distribution.

```{r}
library(EnvStats)
data = c(12.43,11.71,14.41,11.05,9.53,11.66,9.33,11.71,14.35,13.81)
varTest(data, alternative = 'greater', sigma.squared = 2.25)
```

## Equality of Variance Test

Test for the equality of variance for samples from two normal populations

-   **H₀**: The variances of all groups are equal.

-   **H₁**: At least one group has a different variance.

```{r}
x = c(1.3,1.5,1.2,1.7,1.3)
y = c(1.6,1.7,1.8,1.6,1.5)
var.test(x,y, alternative = 'two.sided')
```

# Linear Regression

-   To study the relationship between response and predictor variable

-   Forecast value of variable interest

Process building model

1.  

```{r}
eruption.lm = lm(eruptions~waiting, data = faithful)
summary(eruption.lm)
```

```{r}
plot(eruptions~waiting, data = faithful, main = 'Plot of Y vs X')
abline(eruption.lm, col = 'red')
```

```{r}
pairs(~stack.loss + Air.Flow + Water.Temp + Acid.Conc., data = stackloss)
```

```{r}
stackloss.lm = lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., 
                  data = stackloss)
summary(stackloss.lm)
```

## Dummy-Variable Model

```{r}
class(mtcars$cyl)
mtcars$CYL = as.factor(mtcars$cyl)
mt.lm = lm(mpg~CYL, data = mtcars)
summary(mt.lm)
```

# Exercise

## Question 1

A fisherman breeds red tilapias in a pond in Sungai Terengganu. He claims that the mean weight of the red tilapias exceeds 1.2 kg. Assume that the weight of the population of red tilapias is approximated normal. A random sample of 5 red tilapias was taken from his pond, and their weights were 1.1, 1.2, 1.3, 1.2 and 1.4 kg, respectively. Test at 𝛼 = 0.05 whether the fisherman’s claim is reasonable.

```{r}
x = c(1.1, 1.2, 1.3, 1.2, 1.4)
tilapia = t.test(x, mu = 1.2, alternative = 'greater')
tilapia

# H_{0}: mean weight <= 1.2kg
# H_{1}: mean weight > 1.2kg
# Since p-value > 0.05, We failed to reject H_{0}, We do not have enough evidence to accept the fisherman's claim
```

## Question 2

For 80 randomly selected men, 30 regularly bicycled to campus; while for 100 randomly selected women, 20 regularly bicycled to campus. Test

-   𝐻0: 𝑝1 = 𝑝2

-   H1: 𝑝1 \> 𝑝2

    at 0.05 level of significance.

```{r}
# H_{0}: p1 = p2
# H_{1}: p1 > p2

prop.test(c(30,20), c(80,100), alternative = 'greater', correct = F)

# p-value < 0.05, We reject H_{0}. Proportion of men bicycled to campus greater than proportion of women bicycled to campus
```

## Question 3

The PlantGrowth dataset are results from an experiment on the growth of plants with different treatments. A researcher wants to know if the average weights of the plants in the three experimental circumstances vary significantly. Perform a test to find out the answer for him.

```{r}
data("PlantGrowth")
str(PlantGrowth)
unique(PlantGrowth$group)

# H_{0}: mean_ctrl = mean_trt1 = mean_trt2
# H_{1}: mean_i != mean_j, for some i and j

res.aov = aov(weight~group, data = PlantGrowth)
summary(res.aov)

# p-value < 0.05, we reject H_{0}, At least one pair of means are not equal
```

## Question 4

Based on the Salaries dataset in the car package, build a simple linear model for the salaries of professors against the years of service, i.e salary as the dependent variable and yrs.service as the independent variable. Then build a multiple linear regression with an additional independent variable of yrs.since.phd and a dummy variable model with both yrs.service and rank as independent variables. Write down the equation for the 3 models and compare them.

```{r}
# To view list of dataset in a desired package
data(package='carData')
library(carData)
data(Salaries)
str(Salaries)
head(Salaries)
unique(Salaries$rank)
unique(Salaries$discipline)

salary.slr = lm(salary ~ yrs.service, data = Salaries)
summary(salary.slr)

# p-value < 0.05, the model is significant, 
# salary = 99974.7 + 779.6*yrs.service + error
# R^2 = 0.1121

salary.mlr = lm(salary ~ yrs.service + yrs.since.phd + rank, data = Salaries)
summary(salary.mlr)

# p-value < 0.05, the model is significant
# salary = 80281.3 - 358.3*yrs.service + 263.5*yrs.since.phd + 13805.9*rankAssocProf + 47208.2*rankProf + error
# R^2 = 
```
