---
title: "Admission"
author: "Hazim Fitri"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
---

# Load Library

```{r}
library(dplyr)
library(mice)
```

# Load Data

```{r}
setwd('C:\\Users\\hazim\\OneDrive - Universiti Kebangsaan Malaysia\\Math-Stat\\Assignment 2\\Graduate Admission')

getwd()

data = read.csv('Admission_Predict_Ver1.1.csv')
```

# Exploratory Data Analysis (EDA)

#### Structure of the data

```{r}
str(data)
```

#### Summary of the data

```{r}
summary(data)
```

#### Top 6 data

```{r}
head(data)
```

#### Check for missing data

```{r}
md.pattern(df)
```

#### Correlogram of the data

```{r}
par(bg = "#EEEEEE")
num_df = df %>% select(-Research)
corrgram(num_df, order = FALSE, lower.panel = panel.pts, upper.panel = panel.cor,
         main = 'Correlogram of Graduate Admission')
```

#### Scatter plot (CGPA VS Chance of Admit)

```{r}
par(bg = "#EEEEEE")
plot(df$CGPA, df$Chance.of.Admit, main = 'CGPA vs. Chance of Admit', 
     xlab = 'CGPA', ylab = 'Chance of Admit', col = '#00ADB5')
```

#### Boxplot of impact of reseach expericne on admission

```{r}
par(bg = "#EEEEEE")
df2 = df %>%
  mutate(Research = factor(Research, labels = c('No', 'Yes')))
boxplot(df2$Chance.of.Admit~df2$Research, col = '#00ADB5', 
        xlab = 'Research Experience', ylab = 'Chance of Admit',
        main = 'Impact of Research Experience on Admission')
```

#### Histogram of the dependent variable

```{r}
par(bg = "#EEEEEE")
hist(df$Chance.of.Admit, col = '#00ADB5', xlab = 'Chance of Admit', 
     main = 'Histogram of Chance of Admit')
```

#### Barplot of the University Rating

```{r}
uni_rate = table(df$University.Rating)
par(bg = "#EEEEEE")
barplot(uni_rate, 
        col = '#00ADB5',
        main = "University Rating",
        xlab = "University Rating", 
        ylab = "Count")
```

#### Barplot of the research status

```{r}
research = table(df$Research)
par(bg = "#EEEEEE")
barplot(research,
        col = '#00ADB5', 
        ylab = 'Research Status',
        xlab = 'Count',
        main = 'Research Status')
```

# Data Pre-processing

Convert Research column to factor as it is a categorical data

```{r}
df = data[,-1] %>%
  mutate(Research = as.factor(Research))
```

Review the structure to make sure Research has been converted to factor

```{r}
str(df)
```

# Modelling

#### Fit data into linear regression model

```{r}
model = lm(Chance.of.Admit~.,
           data = df)

summary(model)
```

#### Plot model

```{r}
par(mfrow=c(2,2))
plot(model)
```

#### Correlation graph for significant variable

```{r}
par(bg = "#EEEEEE")
num_df = df %>% select(-Research, -SOP, -University.Rating0)
pairs(num_df, col = '#00ADB5', main = 'Correlation')
```

```{r}
library(car)
vif(model)
```

#### linear regression plot

```{r}
par(bg = "#EEEEEE")

df$Predicted <- predict(model, df)

# Plot Actual vs. Predicted
plot(df$Chance.of.Admit, df$Predicted, 
     xlab = "Actual Chance of Admit", ylab = "Predicted Chance of Admit",
     main = "Actual vs. Predicted Values", col = "#00ADB5", pch = 16)
abline(0, 1, col = "red", lwd = 3)

```
