---
title: "Data Exploration with R"
author: "Hazim Fitri"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
---

# Import/Export Data

## Import data into R

### From text file

```{r}
read.table(file.choose, header = T, sep = '/t')
```

### From csv file

```{r}
read.csv(file.choose)
```

### From excel file

```{r}
library(readxl)
read_excel(file.choose(), sheet = 'Sheet1')

# To find out what sheets are available
excel_sheets(file.choose())
```

### From clipboard

```{r}
read.table(file = 'clipboard', header = T, sep = '/t')
```

## Export data from R

### Save to text file

```{r}
write.table(variable, file = 'Rclass.txt', col.names = T, row.names = F)
```

### Save to csv file

```{r}
write.csv(variable, file = 'Rclass.csv', col.names = T, row.names = F)
```

### Save to excel file

```{r}
library(writexl)
write_xlsx(variable, 'Rclass.xlsx')
```

### Save to clipboard

```{r}
write.table(variable, col.names = T, row.names = F, file = 'clipboard', 
            sep = '/t')
```

# Descriptive Statistics Using R

## Central Tendency

### Mean

```{r}
mean()
rowMeans()
colMeans()
```

### Median

```{r}
median()
```

### Mode

```{r}
x = c(1,2,2,2,3,4,5,6,7,8,8,9)
temp = table(x)
names(temp)[temp == max(temp)]
```

## Quartiles

```{r}
quantile(x)
```

## Variation

### Range

```{r}
# Give 1 output(max - min)
max(x) - min(x)

# Give 2 outputs(min and max)
range(x)
```

### Interquartile Range (IQR)

```{r}
IQR(x)
```

### Variance

```{r}
# Sample variance
var(x)
```

### Standard Deviation

```{r}
#Sample standard deviation
sd(x)
```

### Coefficient of Variation

```{r}
sd(x) / mean(x) * 100
```

## Shape

### Skewness

```{r}

```

### Kurtosis

# Simple Data Wrangling with R

## head, tail, which Function

## subset, sort, order Function

merge

factor

## table

```{r}
wfact <- cut(women$weight,breaks=3)
wtab <- table(wfact)
wtab
```

```{r}
wfact <- cut(women$weight,breaks=3,labels=c("Low","Medium","High"))
table(wfact)
```

```{r}
wfact <- cut(women$weight,breaks=c(100,120,140,160,180))
table(wfact)
```

## sapply, lapply

```{r}
data(rock)
lapply(rock, FUN=mean) # same as colMeans but you'll be able to call each column like a data frame
sapply(rock, FUN=mean) # exactly same as colMeans

```

## apply, do.call

`do.call` allows you to call any R function, but instead of writing out the arguments one by one, you can use a list to hold the arguments of the function.

```{r}
do.call(rbind,lapply(rock,FUN=mean))
```

## is.na, na.omit

## paste

```{r}
mine<-c("a","b","c")
paste(mine,"!")
paste(mine,1:3)
paste0(mine,"!")
paste(mine,1:3,sep=",")
paste(mine,1:3,collapse=",")
```

# Exercise

## Question 1
